Forecasting the 2024 U.S. Presidential Election Using a Model-Based Approach

### 2024-11-02 paper/paper.qmd

library(tidyverse)
library(arrow)
library(dplyr)
library(ggplot2)
```

@@

### The support percentage of Kamala Harris.{#sec-harris-support}

@@

# Model{#sec-model}

@@

The model predicts Harris's support percentage using the following
predictor variables:

@@

@@

Where:

@@

- **Recency Weight**: Applied through an exponential decay function, exp(-recency * 0.1), where "recency" represents the days since the poll was conducted. This ensures that more recent data carries more weight, while older polls have reduced influence.

@@

Finally, to ensure the robustness and accuracy of our model, we conduct
model validation and diagnostics, which are detailed in Appendix B.
These procedures help confirm that the model assumptions are met and
that the predictions are reliable across different polling scenarios.

@@

## Model Results and Interpretation {#sec-result-explain}

The linear regression models constructed for predicting support for Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election utilized both unweighted and weighted configurations. As shown in **Table 1** and **Table 2**, the intercepts for Harris and Trump, estimated at 45.10 and 59.28 respectively in the unweighted models, represent their baseline support levels when all predictors are at reference values. The high \( R^2 \) values of 0.738 for Harris’s unweighted model and 0.860 for her weighted model, along with similar values of 0.635 and 0.890 for Trump, indicate that a significant portion of the variation in support can be explained by the model's predictors. The relatively low Root Mean Square Error (RMSE) values further suggest that the models provide reasonably accurate predictions of support levels.

The results underscore the significance of **national_poll** and **pollster** as key predictors for both candidates' support, highlighting how national trends and pollster differences impact the reported support levels. The **population** predictor further reveals that demographic and regional characteristics contribute meaningfully to support variations, especially in battleground states. These findings enhance our understanding of the dynamics influencing voter sentiment, with methodological and regional factors proving essential in shaping support for each candidate.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of the Linear Regression Model for Predicting Harris' Voting Outcomes by polling results"
#| label: tbl-summary1
modelsummary::msummary(
  list(
    "Harris Unweighted" = harris_unweighted_model, 
    "Harris Weighted" = harris_weighted_model
  ), 
  options = list("longtable" = TRUE))
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of the Linear Regression Model for Predicting Trump's Voting Outcomes by polling results"
#| label: tbl-summary2
modelsummary::msummary(
  list(
    "Trump Unweighted" = trump_unweighted_model,
    "Trump Weighted" = trump_weighted_model
  ), 
  options = list("longtable" = TRUE))
```

@@

@fig-support-trend1 and @fig-support-trend2 illustrate the predicted support for Kamala Harris and Donald Trump from August 1, 2024, to October 27, 2024, using both unweighted and weighted linear regression models. In the unweighted model, we see that Trump's predicted support generally hovers between 45% and 50%, while Harris's support ranges from 35% to 45%. Both candidates' support levels appear relatively stable across the polling period, with minor fluctuations. This model, which treats all polls equally, suggests a consistent lead for Trump, albeit within a fairly close range, as shown by the overlapping grey points representing individual poll results.

The weighted model refines these predictions by applying weights based on recency and sample size, which reduces the influence of older or smaller polls. This approach results in smoother trend lines for both candidates, with less apparent variability compared to the unweighted model. Trump's support remains around the mid-40% range, while Harris’s support stays in the high 30% range. The weighting emphasizes more recent and larger polls, likely providing a more accurate reflection of current public opinion. Overall, both models consistently indicate a lead for Trump, with the weighted model slightly stabilizing the predictions.

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| 
set.seed(233)
####### Prediction for Harris ######
new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 2000
  ),
  state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 2000, replace = TRUE)), 
  pollster = factor(sample(levels(harris_data$pollster), size = 2000, replace = TRUE), 
                    levels = levels(harris_data$pollster)), 
  population = factor(sample(levels(harris_data$population), size = 2000, replace = TRUE), 
                    levels = levels(harris_data$population)),
  sample_size = sample(1:2300, size = 2000, replace = TRUE)
)|>
  mutate(
    national_poll = if_else(state == "National", 1, 0),
    recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),
    combined_weight = recency_weight * sample_size_weight,
    candidate="Kamala Harris"
  )

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
# Predictions for Harris
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(harris_unweighted_model, newdata = new_data_harris),
    predicted_weighted = predict(harris_weighted_model, newdata = new_data_harris)
  )

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Predicted Support Trends for Kamala Harris and Donald Trump using unweighted fitted Linear Model"
#| label: fig-support-trend1
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 1)+
  labs(title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using unweighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Linear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Predicted Support Trends for Kamala Harris and Donald Trump using weighted fitted Linear Model"
#| label: fig-support-trend2
# Plotting with the filtered data
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 1)+
  labs(title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using weighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red"))  +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))
```

@@

@fig-support-state and @fig-support-state2 compare the predicted support for Kamala Harris and Donald Trump in key battleground states (Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin) for the 2024 U.S. election. @fig-support-state uses an unweighted multiple linear regression model, while @fig-support-state2 applies a weighted model, giving greater influence to polls with higher weights based on recency and sample size. In both figures, Trump’s support, represented by red lines, and Harris’s support, shown in blue, show close competition across most states, with slight variations in trends across the polling period from August to October 2024.

The weighted model in @fig-support-state2 introduces subtle shifts, with larger circles indicating polls with greater influence on the predicted trend. This approach captures more recent or higher-quality polls, offering a potentially more accurate representation of voter sentiment shifts over time. The close alignment of support lines for both candidates suggests tight races in these states, which are crucial for the election outcome. The use of weighted and unweighted models highlights the importance of data weighting in enhancing the reliability of electoral forecasts in dynamic political environments.
```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Predicted Support Trends for Kamala Harris and Donald Trump by State using Fitted Linear Model"
#| label: fig-support-state
## unweighted
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 1)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump by State \n in 2024 US Election (From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using unweighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       caption = "Linear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Predicted Support Trends for Kamala Harris and Donald Trump by State using Weighted Fitted Linear Model"
#| label: fig-support-state2
## weighetd
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 1)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump by State \n in 2024 US Election (From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using weighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Linear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
```

@@

@fig-support-pollster1 and @fig-support-pollster2 depict the predicted support for Kamala Harris and Donald Trump in the 2024 U.S. election, from August 1, 2024, to October 27, 2024, using a weighted multiple linear regression model. The red line shows fluctuations in Trump’s support, generally between 40% and 55%, with some noticeable peaks and dips, indicating variability over time. In contrast, Harris’s support, represented by the blue line, remains more stable, consistently ranging around 40% to 50%, suggesting less volatility in voter sentiment for her.

Pollsters such as AtlasIntel, Emerson, Beacon/Shaw, Siena/NY, and YouGov are represented by different colors, with circle sizes reflecting weights based on factors like recency and sample size. Larger circles correspond to higher-weighted polls, indicating they had a greater influence on the predictions. The model incorporates national poll data, adjustments for pollster effects, and population considerations, allowing for a refined prediction approach. This approach ensures that more recent and larger polls play a more significant role, capturing shifts in voter sentiment as the election approaches.
```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Predicted Support Trends for Kamala Harris and Donald Trump by Pollster using Unweighted Fitted Linear Model"
#| label: fig-support-pollster1

# Calculate the top 10 pollsters by frequency in the data
top_pollsters <- combined_data_total %>%
  count(pollster) %>%
  top_n(5, n) %>%
  pull(pollster)

@@

ggplot(combined_data_total, aes(x = end_date, y = pct, size = combined_weight)) +
  geom_point(aes(fill = pollster_group), alpha = 0.5, shape = 21, color = "black") + 
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Kamala Harris"), 
            aes(x = end_date, y = predicted_unweighted), color = "blue", size = 1) +
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Donald Trump"), 
            aes(x = end_date, y = predicted_unweighted), color = "red", size = 1) +
  facet_wrap(~ candidate, ncol = 2) +
  labs(
    title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election by Top 5 Pollster \n(From August 1, 2024 to October 27, 2024)",
    subtitle = "predicted using weighted multiple linear regression model",
    x = "End Date of the Poll",
    y = "Predicted Support (%)",
    size = "Weight",
    fill = "Pollster",
    caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage = national_poll + pollster + population"
  ) +
  scale_fill_manual(values = c(setNames(rep("grey", length(unique(combined_data_total$pollster_group))), "Other"),
                               setNames(rainbow(10), top_pollsters))) + # Assign colors to top pollsters and grey to others
  theme_classic() +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",  # Place legends in a horizontal row
    legend.title.align = 0.5,   # Center-align legend titles for better appearance
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(colour = "black", fill = "white")
  ) +
  guides(
    size = guide_legend(order = 1, title.position = "top", label.position = "left"),
    fill = guide_legend(order = 2)
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Predicted Support Trends for Kamala Harris and Donald Trump by Pollster using Weighted Fitted Linear Model"
#| label: fig-support-pollster2

ggplot(combined_data_total, aes(x = end_date, y = pct, size = combined_weight)) +
  geom_point(aes(fill = pollster_group), alpha = 0.5, shape = 21, color = "black") + 
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Kamala Harris"), 
            aes(x = end_date, y = predicted_weighted), color = "blue", size = 1) +
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Donald Trump"), 
            aes(x = end_date, y = predicted_weighted), color = "red", size = 1) +
  facet_wrap(~ candidate, ncol = 2) +
  labs(
    title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election by Top 5 Pollster \n(From August 1, 2024 to October 27, 2024)",
    subtitle = "predicted using weighted multiple linear regression model",
    x = "End Date of the Poll",
    y = "Predicted Support (%)",
    size = "Weight",
    fill = "Pollster",
    caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage = national_poll + pollster + population"
  ) +
  scale_fill_manual(values = c(setNames(rep("grey", length(unique(combined_data_total$pollster_group))), "Other"),
                               setNames(rainbow(10), top_pollsters))) + # Assign colors to top pollsters and grey to others
  theme_classic() +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",  # Place legends in a horizontal row
    legend.title.align = 0.5,   # Center-align legend titles for better appearance
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(colour = "black", fill = "white")
  ) +
  guides(
    size = guide_legend(order = 1, title.position = "top", label.position = "left"),
    fill = guide_legend(order = 2)
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

@@

# Additional data details {#appendix-a}

@@

# Model details {#appendix-b}

@@

# Residual plot for unweighted model
plot(harris_unweighted_model, which = 1, main = "Residual Plot of Unweighted Linear Regression Model")

# Q-Q plot for the residuals of the unweighted model
qqnorm(residuals(harris_unweighted_model), main = "Q-Q Plot of Unweighted Linear Regression Model Residuals")
qqline(residuals(harris_unweighted_model), col = "red")

# Residual plot for weighted model
plot(harris_weighted_model, which = 1, main = "Residual Plot of Weighted Linear Regression Model")

# Q-Q plot for the residuals of the weighted model
qqnorm(residuals(harris_weighted_model), main = "Q-Q Plot of Weighted Linear Regression Model Residuals")
qqline(residuals(harris_weighted_model), col = "red")

@@

## Survey Questions

>>>>

library(tidyverse)
library(arrow)
library(dplyr)
library(ggplot2)
library(modelsummary)
library(RColorBrewer)
library(gt)
library(future)
```

@@

### The support percentage of Kamala Harris. {#sec-harris-support}

@@

# Model {#sec-model}

@@

The model predicts Harris's support percentage by constructing multiple
linear regression model using the following predictor variables:

@@

We include Pollster and State variables as controls to account for potential biases introduced by different polling organizations and regional variations in voter preferences.

@@

Where: $W$ is the weight parameters calculated from the product of
**sample size weight** and recency **weight**.

@@

-   **Recency Weight**: Applied through an exponential decay function,
    exp(-recency \* 0.1), where "recency" represents the days since the
    poll was conducted. This ensures that more recent data carries more
    weight, while older polls have reduced influence.

@@

Finally, to ensure the robustness and accuracy of our model, we conduct
model validation and diagnostics, which are detailed in Appendix B. We
use K-Fold Cross-Validation to validate the model, use Q-Q Plot and
Residuals vs fitted diagram to diagnostic the model. These procedures
help confirm that the model assumptions are met and that the predictions
are reliable across different polling scenarios.

@@

## Model Results

The linear regression models constructed for predicting support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election
utilized both unweighted and weighted configurations.The model summary of support for Harris is shown in @tbl-summary1, support for Trump is shown in @tbl-summary2. As shown in
@tbl-summary1 and @tbl-summary2, the intercepts for Harris and Trump,
estimated at 45.10 and 59.28 respectively in the unweighted models,
represent their baseline support levels when all predictors are at
reference values. The high ( R\^2 ) values of 0.738 for Harris’s
unweighted model and 0.860 for her weighted model, along with similar
values of 0.635 and 0.890 for Trump, indicate that a significant portion
of the variation in support can be explained by the model's predictors.
The relatively low Root Mean Square Error (RMSE) values further suggest
that the models provide reasonably accurate predictions of support
levels.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Harris Support Based on Polling Data"
#| label: tbl-summary1
msummary(
  list("Unweighted" = harris_unweighted_model, "Weighted" = harris_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska, the reference level of population is likely voter, and the reference level of pollsters is Alaska is Angus Reid.")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Trump Support Based on Polling Data"
#| label: tbl-summary2
msummary(
  list("Unweighted" = trump_unweighted_model, "Weighted" = trump_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska,the reference level of population is likely voter, and the reference level of pollsters is Alaska is ABC/Washington Post.")
```

The results underscore the significance of **national_poll** as key predictors for both candidates' support, highlighting how national trends and pollster differences impact the
reported support levels, as polls from the National level increase the
support of Harris by 3.262%, decrease the support for Trump by 7.943%.
The **population** predictor indicates that the type of respondent group
contributes meaningfully to variations in support. With "likely voters"
as the baseline level, the coefficient for "registered voters"
(populationrv) suggests a lower level of support for both Harris and
Trump in this group compared to likely voters. Specifically, the support
of registered voters is 2.1095 (unweighted model) and 1.0322 (weighted
model) lower than the likely voters. This implies that, on average,
support among registered voters is lower than that among likely voters.

@@

@fig-support-trend illustrate the predicted support for Kamala Harris
and Donald Trump from August 1, 2024, to October 27, 2024, using both
unweighted and weighted linear regression models. In the
@fig-support-trend-1 unweighted model, we see that Trump's predicted
support generally hovers between 45% and 50%, while Harris's support
ranges from 35% to 45%. Both candidates' support levels appear
relatively stable across the polling period, with minor fluctuations.
This model, which treats all polls equally, suggests a consistent lead
for Trump, albeit within a fairly close range, as shown by the
overlapping grey points representing individual poll results.

The weighted model @fig-support-trend-2 refines these predictions by
applying weights based on recency and sample size, which reduces the
influence of older or smaller polls. This approach results in smoother
trend lines for both candidates, with less apparent variability compared
to the unweighted model @fig-support-trend-1. Trump's support remains
around the mid-40% range, while Harris’s support stays in the high 30%
range. The weighting emphasizes more recent and larger polls, likely
providing a more accurate reflection of current public opinion. Overall,
both models consistently indicate a lead for Trump, with the weighted
model slightly stabilizing the predictions.

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

set.seed(233)
plan(multisession)
####### Prediction for Harris ######
new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 2000
  ),
  state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 2000, replace = TRUE)), 
  pollster = factor(sample(levels(harris_data$pollster), size = 2000, replace = TRUE), 
                    levels = levels(harris_data$pollster)), 
  population = factor(sample(levels(harris_data$population), size = 2000, replace = TRUE), 
                    levels = levels(harris_data$population)),
  sample_size = sample(1:2300, size = 2000, replace = TRUE)
)|>
  mutate(
    national_poll = if_else(state == "National", 1, 0),
    recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),
    combined_weight = recency_weight * sample_size_weight,
    candidate="Kamala Harris"
  )

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
# Predictions for Harris
plan(multisession)
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(harris_unweighted_model, newdata = new_data_harris),
    predicted_weighted = predict(harris_weighted_model, newdata = new_data_harris)
  )

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Harris Kamala leads Donald Trump an average 5% in Predicted Support in the 2024 US Election (From August 1, 2024, to October 27, 2024)"
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 2
#| label: fig-support-trend
# Plotting with the filtered data
plan(multisession)
##unweighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 1)+
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Linear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
##weighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 1)+
  labs(x = "End Date of the Poll",
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red"))  +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))
```

@@

@fig-support-state compare the predicted support for Kamala Harris and
Donald Trump in key battleground states (Arizona, Georgia, Michigan,
Nevada, North Carolina, Pennsylvania, and Wisconsin) for the 2024 U.S.
election. @fig-support-state-1 uses an unweighted multiple linear
regression model, while @fig-support-state-2 applies a weighted model,
giving greater influence to polls with higher weights based on recency
and sample size. In both figures, Trump’s support, represented by red
lines, and Harris’s support, shown in blue, show close competition
across most states, with slight variations in trends across the polling
period from August to October 2024.

In the @fig-support-state-1 , support for Kamala Harris and Donald Trump
fluctuates widely across battleground states, with Harris holding slight
leads in Pennsylvania and North Carolina but facing tight competition
overall. Without weighting, polls contribute equally, creating more
volatility.

The @fig-support-state-2 stabilizes these trends by emphasizing recent,
larger polls. Harris's support becomes more consistent, particularly in
Pennsylvania, North Carolina, and Nevada, where she shows a clearer
lead. This weighted approach provides a more refined snapshot,
highlighting Harris’s likely advantage in these key states.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 2
#| fig-cap: "Harris is likely to lead Trump in Pennsylvania, North Carolina, and Nevada by Model Prediction"
#| label: fig-support-state
## unweighted
plan(multisession)
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)

ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 0.5)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       caption = "Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

## weighetd
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 0.5)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight\nCandidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
```

@@

@fig-support-pollster depict the predicted support for Kamala Harris and
Donald Trump in the 2024 U.S. election, from August 1, 2024, to October
27, 2024, using a weighted multiple linear regression model. The red
line shows fluctuations in Trump’s support, generally between 40% and
55%, with some noticeable peaks and dips, indicating variability over
time. In contrast, Harris’s support, represented by the blue line,
remains more stable, consistently ranging around 40% to 50%, suggesting
less volatility in voter sentiment for her.

Major Pollsters are represented by different colors, with circle sizes
reflecting weights based on factors like recency and sample size. Larger
circles correspond to higher-weighted polls, indicating they had a
greater influence on the predictions. The model incorporates national
poll data, adjustments for pollster effects, and population
considerations, allowing for a refined prediction approach. This
approach ensures that more recent and larger polls play a more
significant role, capturing shifts in voter sentiment as the election
approaches.

Siena/NYT and Quinnipiac consistently show stronger support for Harris,
with Siena/NYT polls placing her around or above 50%. In contrast,
Emerson polls display a closer race, often showing near-equal support
for both candidates. AtlasIntel and Beacon/Shaw lean toward Trump, with
AtlasIntel showing more volatility and Beacon/Shaw stabilizing near
45-50% for Trump. Polls from other organizations, in grey, introduce
variability but hold less influence in the model. Overall, Siena/NYT and
Quinnipiac favor Harris, while Emerson, AtlasIntel, and Beacon/Shaw
present a more competitive picture.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 2
#| fig-cap: "Major Pollsters Siena/NYT and Quinnipiac favor Harris, while Emerson, AtlasIntel, and Beacon/Shaw present a more competitive picture"
#| label: fig-support-pollster
plan(multisession)
# Calculate the top 10 pollsters by frequency in the data
top_pollsters <- combined_data_total %>%
  count(pollster) %>%
  top_n(5, n) %>%
  pull(pollster)

@@

ggplot(combined_data_total, aes(x = end_date, y = pct)) +
  geom_point(aes(fill = pollster_group), alpha = 0.7, shape = 21) + 
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Kamala Harris"), 
            aes(x = end_date, y = predicted_unweighted), color = "blue", size = 1) +
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Donald Trump"), 
            aes(x = end_date, y = predicted_unweighted), color = "red", size = 1) +
  labs(
    x = "End Date of the Poll",
    y = "Predicted Support (%)",
    color = "Candidate",
    fill = "Top 5 Pollster",
    caption = "Candidate Support Percentage = national_poll + pollster + population + state"
  )+ # Smoothed trend line
  facet_wrap(~ candidate,ncol = 2) +
  scale_fill_manual(values = c(setNames(rep("grey", length(unique(combined_data_total$pollster_group))), "Other"),
                               setNames(brewer.pal(10, "Set1"), top_pollsters))) + 
  theme_classic() +
  theme(
    legend.position = "right",
    legend.box = "vertical",  # Place legends in a horizontal row
    legend.title.align = 0.5,   # Center-align legend titles for better appearance
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(colour = "black", fill = "white")
  ) +
  guides(
    size = guide_legend(order = 1, title.position = "top", label.position = "left"),
    fill = guide_legend(order = 2)
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

ggplot(combined_data_total, aes(x = end_date, y = pct, size = combined_weight)) +
  geom_point(aes(fill = pollster_group), alpha = 0.7, shape = 21) + 
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Kamala Harris"), 
            aes(x = end_date, y = predicted_weighted), color = "blue", size = 1) +
  geom_line(data = combined_predictions_filtered %>% filter(candidate == "Donald Trump"), 
            aes(x = end_date, y = predicted_weighted), color = "red", size = 1)+
  labs(
    x = "End Date of the Poll",
    y = "Predicted Support (%)",
    size = "Weight",
    color = "Candidate",
    fill = "Top 5 Pollster",
    caption = "Weights: recency_weight * sample_size_weight \nCandidate Support Percentage = national_poll + pollster + population + state"
  )+ # Smoothed trend line
  facet_wrap(~ candidate,ncol = 2) +
  scale_fill_manual(values = c(setNames(rep("grey", length(unique(combined_data_total$pollster_group))), "Other"),
                               setNames(brewer.pal(10, "Set1"), top_pollsters))) +
  theme_classic() +
  theme(
    legend.position = "right",
    legend.box = "vertical",  # Place legends in a horizontal row
    legend.title.align = 0.5,   # Center-align legend titles for better appearance
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(colour = "black", fill = "white")
  ) +
  guides(
    size = guide_legend(order = 1, title.position = "top", label.position = "left"),
    fill = guide_legend(order = 2)
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

@@

# Additional data details {#first-data-details}

@@

# Model details {#sec-model-details}

@@

# Residual plot for unweighted model
plot(harris_unweighted_model, which = 1)

# Q-Q plot for the residuals of the unweighted model
qqnorm(residuals(harris_unweighted_model) )
qqline(residuals(harris_unweighted_model), col = "red")

# Residual plot for weighted model
plot(harris_weighted_model, which = 1)

# Q-Q plot for the residuals of the weighted model
qqnorm(residuals(harris_weighted_model))
qqline(residuals(harris_weighted_model), col = "red")

@@
